\sysfull

### 2017-07-31 paper/background.tex
added some edits to section 2.  still needs  alot, especially in the task section

\sys addresses the problem of intermittently executing task-based computations on energy-harvesting hardware, without being limited by inflexible, static task decomposition and without being limited to fully non-volatile main memory. In this section, we provide background on energy harvesting systems and the intermittent software execution model. We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.

@@

Energy harvesting devices are systems that operate using energy extracted from their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar energy. By operating entirely using environmental energy, energy-harvesting devices need not use a tethered power supply or battery. Instead, devices typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.

Batteryless operation has a number of important advantages, making intermittent computing an important research domain. Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable. The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}. Batteries constitute a potential environmental hazard because not all batteries are recycled. Batteries are fragile and limited in their number of charge/discharge cycles, requiring costly maintenance that is possibly only through physical device access, which may be difficult if devices are deployed in inaccessible places, e.g. in space~\cite{kicksat}. By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently: operating discharge time---is orders of magnitude less than a battery. \todo{Compress text}{Brandon}

Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.

@@

The biggest technical problem with energy provided from the ambient to battery-less platforms is that harvested energy is not stable and difficult to predict accurately. Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation. Energy breaks happen every hundred of milliseconds, refer again to e.g.~\cite[Fig. 1]{mementos}. This causes each program running on such platform to restart and requires a fundamentally new model for program execution: \emph{intermittent} (discontinuous) execution model. In the conventional (or \emph{continuous}) execution model a power failure is seen as an \emph{exception} that may reset the progress of a program to its start. Whereas, in an intermittent execution model a power failure is regarded as a temporary \emph{pause} to the execution consequently prolonging execution time. A conceptual implementation of intermittent execution model is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.

\subsection{Intermittent Computing}
\label{sec:background_consistency}

Naturally, content of a volatile memory (i.e. run-time stack, global and local variables) and program registers of the microcontroller are erased after each power failure. Only what resides in a non-volatile memory, e.g. FRAM, persists after system reboot. A solution to sustain the progress of the program from the last power failure, instead from the beginning (also denoted as the \emph{Sisyphus} effect~\cite[Sec. 2]{mementos}) is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}. The periodicity, and a position of a checkpoint, as well as the way to combat the \emph{program state consistency} phenomenon (discussed extensively in e.g.~\cite{chain,alpaca}), is the core of the contribution behind each of the proposed checkpoining solution. For the completeness of this exposition we will provide our own example demonstrating this phenomenon, viz. please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec. 2.1]{chain}.

@@

Let us take an example of a fast procedure to calculate cyclic redundancy check (CRC) from~\cite{hicks_mibench2_2016}. Therein a CRC must be calculated for $\texttt{nBytes}$-long message $\texttt{msg}$ with specified initial reminder $\texttt{Rmnd}$ and constant, $\texttt{Rmnd}$ variable type-dependent, $\texttt{Wdth}$. Despite regular, fixed position checkpoints (line 3 in Fig.~\ref{fig:crc_example}) of all variables which are declared as non-volatile (denoted as \texttt{NV}) variable $\texttt{data}$ is calculated incorrectly. This is because of a write after read (WAR) hazard of variable $\texttt{Rmnd}$. WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more specifically---updated before a new execution loop.

\subsection{(Task-based) Intermittent Computing: Background}

Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
)
\subsubsection{Baseline---Sequential Execution Model}

@@

\subsubsection{Alternative---Modular Execution Model}

>>>>

\In this section, we provide background on energy harvesting systems and the
intermittent software execution model. We then discuss the key limitations of
existing task-based execution models for intermittent computing that \sys
addresses.

@@

Energy harvesting devices operate using energy extracted from their
environment, from sources such as radio waves (RF)~\cite{wisp} and solar
energy. Energy-harvesting devices need not use a tethered power supply or
battery and instead, typically collect energy into a (super-)capacitor, operate
when sufficient energy has accumulated, and upon depleting the stored energy,
turn off and wait to collect more energy. 
%
Batteryless operation has a number of important advantages, making intermittent
computing an important research domain. Supplying power to
billions~\cite{gartner_iot} of embedded computers using batteries is not
sustainable. The European Commission estimates that more than 160 kilotons of
consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
Batteries are an environmental risk, are fragile, are limited in their number
of charge/discharge cycles, and may require costly physical maintenance that is
difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast,
super-capacitors are durable, promising millions of charge/discharge
cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to
capacitor-based energy storage is that capacitor energy density---and
consequently operating discharge time---is orders of magnitude less than a
battery.

%Given current technology development, battery-less systems are best suited for
%very long-term sensing and monitoring where access to recharge is either
%prohibitive or impossible. These include battery-less image capture and
%processing~\cite{naderiparizi_rfid_2015}, animal
%monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
%and digestible~\cite{nadeau_naturebio_2017} sensors.

@@

\subsection{Intermittent Computing}
\label{sec:background_consistency}
Energy-harvesting devices execute software according to the {\em intermittent
execution model}~\cite{dino,chain,alpaca,ratchet}.  Physically, a device
charges until a threshold, operates briefly until its energy is depleted, shuts
down recharges, and repeats the cycle.  Software running on the device executes
{\em intermittently} because energy is only available between when the
capacitor reaches its threshold and when it is depleted. An intermittent
execution is composed of operating periods interspersed with power failures.
The frequency of power failures in an intermittent execution depends on the
size of the device's energy storage buffer: a larger buffer allows longer
operating periods.  Energy-harvesters provide input power that is orders of
magnitude less than a device's operating power, making recharging negligible
during operation.

Software has different behavior when executed intermittently than when executed
with continuous power.  Each power failure clears volatile memory (registers,
stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
power failures. On a power failure, control flows back to a prior point in the
execution. By default, after a power failure control flows to the beginning of
{\tt main()}; a host of recent work proposed variations of the intermittent
execution model that restart execution differently. Some systems restore to a
checkpoint of execution context (i.e., registers and stack) stored in
non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.

@@

Data may become inconsistent if an attempt to execute some computation {\em
writes} to a non-volatile variable, then power fails, then a second attempt to
re-execute the same computation {\em reads} the value written in the first
attempt.  Figure~\ref{fig:code_demo_incosistency} illustrates how state can
become inconsistent in an intermittent execution using cyclic redundancy check
(CRC) code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for
an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder
$\texttt{Rmnd}$. Even with programmer defined checkpoints (i.e., line 3 in
Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables
(labeled \texttt{NV}), the code may compute  $\texttt{data}$ incorrectly.  The
code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.  WAR
hazard is caused by consecutive write and read in-between two consecutive
checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is
written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war})
$\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
specifically---updated before a new execution loop.

As the figure illustrates, checkpoint-based systems risk violating the
consistency of volatile and non-volatile memory~\cite{dino}. To maintain memory
consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of
non-volatile variables with the checkpointed execution context. Task-based
systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail
next, ensure state is consistent by using programmer, compiler, and runtime
support.

\subsection{Task-based Intermittent Computing}
\TODO{This sequential / modular split seems pretty odd to me.  I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.  Anything with static (programmer-written or compiler inserted) boundaries applies}
Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
)

@@
